Visualize Stan Mixed Models

### 2021-03-22 Stan/Stan Base Models.Rmd

```{r setup, include=FALSE}
library(rstanarm)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
library("posterior")
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

@@

plot(density(posterior_flow$lambda), main = "Lambda")
abline(v = lm_lambda, col = 4, lty = 2)
```
####Abort
```{r abort model}
y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
vol = flower$volume_t
N = nrow(flower)
stan_data <- list(N = N, ## number of observations
                  y = y, ## response survival next year
                  vol = vol
                  )

@@

data {
 int < lower = 1 > N; // Sample size
 int <lower = 0, upper = 1> y[N]; // Outcome
 vector[N] vol;	//size_t
}
parameters {
  real beta0; //intercept, unique to ant sp
  real beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}
model {
 // Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(beta0 + beta1 * vol);
}
",

"abort_lm1.stan")
## Check that you wrote a file
stanc("abort_lm1.stan")
## Save the file path
stanny <- "abort_lm1.stan"
## Now we fit the model
fit_abort <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_abort

posterior_abort <- as.data.frame(fit_abort)
print(colnames(posterior_abort))
```

@@

"grow_ant_lm1.stan")
## Check that you wrote a file
stanc("grow_ant_lm1.stan")
## Save the file path
#stanny <- "grow_ant_lm1.stan"
## Now we fit the model
fit_grow_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_grow_ant
#class(fit)
posterior_grow_ant <- as.data.frame(fit_grow_ant)
print(colnames(posterior_grow_ant))
lm_alpha_other <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(grow_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(grow_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(grow_model_ant_interact)$coeff[5]
lm_beta_other <- summary(grow_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(grow_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(grow_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(grow_model_ant_interact)$coeff[8]
lm_sigma <- sigma(grow_model_ant_interact)  # the residual error

@@

####Abortion Model

```{r abortion model}
y <- cbind(flower$Goodbuds_t, flower$ABFlowerbuds_t)
vol = flower$volume_t
N = nrow(flower)
N_ant = 4
ant = flower$ant
stan_data <- list(N = N, ## number of observations
                  y = y, ## response survival next year
                  vol = vol,
                  N_ant = N_ant,
                  ant = ant
)

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i];
  }
}
model {
 // Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100); 
 // Model
 y ~ bernoulli_logit(mu);
}
",

"abort_ant_lm1.stan")
## Check that you wrote a file
stanc("abort_ant_lm1.stan")
## Save the file path
stanny <- "abort_ant_lm1.stan"
## Now we fit the model
fit_abort_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_abort_ant

posterior_abort_ant <- as.data.frame(fit_abort_ant)
print(colnames(posterior_abort_ant))

lm_alpha_other <- summary(abort_model_ant)$coeff[1]  # the intercept
lm_alpha_crem <- summary(abort_model_ant)$coeff[3]
lm_alpha_liom <- summary(abort_model_ant)$coeff[4]
lm_alpha_vacant <- summary(abort_model_ant)$coeff[5]
lm_beta_other <- summary(abort_model_ant)$coeff[2]  # the slope
lm_beta_crem <- summary(abort_model_ant)$coeff[6]
lm_beta_liom <- summary(abort_model_ant)$coeff[7]
lm_beta_vacant <- summary(abort_model_ant)$coeff[8]

par(mfrow = c(2,4))
#Plot Intercepts
plot(density(posterior_abort_ant$`beta0[1]`), main = "Intercept Other")
abline(v = lm_alpha_other, col = 4, lty = 2)
plot(density(posterior_abort_ant$`beta0[2]`), main = "Intercept Crem")
abline(v = lm_alpha_crem, col = 4, lty = 2)
plot(density(posterior_abort_ant$`beta0[3]`), main = "Intercept Liom")
abline(v = lm_alpha_liom, col = 4, lty = 2)
plot(density(posterior_abort_ant$`beta0[4]`), main = "Intercept Vacant")
abline(v = lm_alpha_vacant, col = 4, lty = 2)
#Plot Slopes
plot(density(posterior_abort_ant$`beta1[1]`), main = "Intercept Other")
abline(v = lm_beta_other, col = 4, lty = 2)
plot(density(posterior_abort_ant$`beta1[2]`), main = "Intercept Crem")
abline(v = lm_beta_crem, col = 4, lty = 2)
plot(density(posterior_abort_ant$`beta1[3]`), main = "Intercept Liom")
abline(v = lm_beta_liom, col = 4, lty = 2)
plot(density(posterior_abort_ant$`beta1[4]`), main = "Intercept Vacant")
abline(v = lm_beta_vacant, col = 4, lty = 2)
```

>>>>

```{r setup, include=FALSE}
library(rstanarm)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
#library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
library("posterior")
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

@@

plot(density(posterior_flow$lambda), main = "Lambda")
abline(v = lm_lambda, col = 4, lty = 2)
```
####Viability
```{r Viability model}
y_good <- flower$Goodbuds_t
y_abort <- flower$AbABFlowerbuds_t
y_tot <- flower$TTotFlowerbuds_t
vol = log(flower$volume_t)
N = nrow(flower)
stan_data <- list(N = N, ## number of observations = trials
                  y_good = y_good, ## good buds = successes
                  y_abort = y_abort, ## aborted buds data
                  y_tot = y_tot, ## number of trials
                  vol = vol
                  )

@@

data {
 int < lower = 1 > N; // Sample size
 int <lower = 0> y_good[N]; // Number of viable seeds
 int<lower=0> y_tot[N]; // number of trials
 vector[N] vol;	//size_t
}
parameters {
  real beta0; //intercept, unique to ant sp
  real beta1; //slope, unique to ant sp
  real<lower = 0> sigma;
  real<lower=0,upper=1> v0;
  real<lower=0> a_v;
}
transformed parameters{
  real<lower=0,upper=1> predV[n_v];
  for(i in 1:N){
    predV[i] = v0 * (1 - pow(SR_v[i],a_v) ) + 0.00001;
  }
}
model {
 // Priors
 beta0 ~ normal(0,100);
 beta1 ~ normal(0,100); 
 // Model
 y ~ binomial_logit(beta0 + beta1 * vol);
}
",

"viab_lm1.stan")
## Check that you wrote a file
stanc("vian_lm1.stan")
## Save the file path
stanny <- "viab_lm1.stan"
## Now we fit the model
fit_viab <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)

posterior_viab <- as.data.frame(fit_viab)
```

@@

"grow_ant_lm1.stan")

## Check that you wrote a file
stanc("grow_ant_lm1.stan")
## Save the file path
stanny <- "grow_ant_lm1.stan"
## Now we fit the model
fit_grow_ant <- stan(file = stanny, data = stan_data, warmup = 50, iter = 100, chains = 1, cores = 2, thin = 1)
fit_grow_ant
#class(fit)
posterior_grow_ant <- as.data.frame(fit_grow_ant)
lm_alpha_other <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(grow_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(grow_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(grow_model_ant_interact)$coeff[5]
lm_beta_other <- summary(grow_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(grow_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(grow_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(grow_model_ant_interact)$coeff[8]
lm_sigma <- sigma(grow_model_ant_interact)  # the residual error

@@

####Viability Model
```{r Viability model}
good <- flower$Goodbuds_t
abort <- flower$ABFlowerbuds_t
tot <- flower$TotFlowerbuds_t
vol = log(flower$volume_t)
N = nrow(flower)
N_ant = 4
ant = flower$ant
stan_data <- list(N = N, ## number of observations = trials
                  N_ant = N_ant,
                  ant = ant,
                  good = good, ## good buds = successes
                  abort = abort, ## aborted buds data
                  tot = tot, ## number of trials
                  vol = vol
)

write("// Stan model for simple total flowers regression
data {
 int < lower = 1 > N; // Sample size
 int <lower = 0> good[N]; // Number of viable seeds
 int <lower = 1> N_ant; // number of ant states
 int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
 int<lower=0> tot[N]; // number of trials
 int<lower=0> abort[N];
 vector[N] vol;	//size_t
}
parameters {
  real <lower = 0, upper = 1> theta[N]; // probability of viability for each bud
  //
  real<lower=0,upper=1> v0;
  real<lower=0> a_v;
}
transformed parameters{
  real<lower=0,upper=1> predV[N]; // proportion viable for each plant?
  // Prediction for seed viability
  for(i in 1:N){
    predV[i] = v0 * pow(theta[ant[i]],good[i]) * (1-theta[ant[i]])^abort[i] ;
  }
}
model {
 // Priors
 v0  ~ beta(10,1);  // intercept viability model
  a_v ~ gamma(1,1);  // decay in viability with SR

  good ~ binomial(tot, predV);
}
", "viab_ant.stan")
## Check that you wrote a file
stanc("viab_ant.stan")
## Save the file path
stanny <- "viab_ant.stan"
## Now we fit the model
fit_viab <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)

posterior_viab_ant <- as.data.frame(fit_viab)

lm_alpha_other <- summary(viab_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(viab_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(viab_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(viab_model_ant_interact)$coeff[5]
lm_beta_other <- summary(viab_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(viab_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(viab_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(viab_model_ant_interact)$coeff[8]

par(mfrow = c(1,2))
#Plot Slopes
plot(density(posterior_viab_ant$`theta[1]`), main = "Viability Slope Distributions", col = "black", lwd = 2, xlim = c(-1,3), ylim =  c(0,2))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_viab_ant$`theta[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_viab_ant$`theta[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_viab_ant$`theta[4]`), col = "pink", lwd = 2)
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
```

### 2021-03-22 Stan/Visualize Stan Mixed.Rmd

## Multinomial
#### 1
```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}
transformed data {
  real<lower = 0> mean_y = mean(to_vector(y));
  real<lower = 0> sd_y = sd(to_vector(y));
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]]  + u[plot[i]] + w[year[i]];
  }
}
model {
  // Model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = multinomail_rng(mu);
  real<lower = 0> mean_y_rep = mean(to_vector(y_rep));
  real<lower = 0> sd_y_rep = sd(to_vector(y_rep));
}
",
"multinomial1.stan")
#stanc("multinomial1.stan")
## Save the file path
stanny <- "multinomial1.stan"
## Now we fit the model
fit_multi1 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 4, cores = 2, thin = 1)
#class(fit)
posterior_multi1 <- as.data.frame(fit_multi1)
```

>>>>

## Multinomial
#### 1
```{r}
N <- nrow(data)
vol = log(data$volume_t)
ant <- data$ant
N_ant <- length(unique(data$ant))
y <- data$ant1
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  N_ant = N_ant, ## number of ants
                  K = K, ## number of covariates (only volume I think)
                  vol = vol, ## Volume in year t
                  y = y, ## ant state in year t1
                  N_Year = N_Year,
                  N_Plot = N_Plot,
                  year = year,
                  plot = plot,
                  ant = ant ## ant state in year t
)
write("
data {
    int<lower=2> N_ant; // of alternatives/outcomes
    int<lower=1> N; // of observations
    int <lower=0,upper=N_ant> y[N];
    int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
    vector[N] vol;
    int<lower=1> N_Year; //number of plots
    int<lower=1> N_Plot; //number of years
    int<lower=1, upper=N_Plot> plot[N]; // plot
    int<lower=1, upper=N_Year> year[N]; // year
}

parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD

}
transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]]  + u[plot[i]] + w[year[i]];
  }
}
model {
  // Model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
    y ~ categorical_logit(mu);
}
generated quantities {
  int<lower = 0> y_rep[N] = multinomial_rng(mu);
}
",
"multinomial1.stan")
#stanc("multinomial1.stan")
## Save the file path
stanny <- "multinomial1.stan"
## Now we fit the model
fit_multi1 <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 4, cores = 2, thin = 1)
#class(fit)
posterior_multi1 <- as.data.frame(fit_multi1)
```
